\documentclass{amsart}
\usepackage[british]{babel}
\usepackage{url}
\usepackage{hyperref}
\usepackage{xcolor}
\usepackage{enumerate}
\definecolor{edblue}{RGB}{0,0,153}
\hypersetup{
  pdftitle={\@title{}},
  colorlinks=true,
  linkcolor=edblue,anchorcolor=edblue,citecolor=edblue,filecolor=edblue,urlcolor=edblue
}%
\usepackage{geometry}
\geometry{a4paper,textwidth=5.5in,textheight=8.5in,marginparsep=7pt,marginparwidth=.6in}
\usepackage{graphicx}
\usepackage{wrapfig}
\title{Report on Scotgov Stuff}
\author{Peter Buneman \and William Waites}
\begin{document}
\maketitle
\tableofcontents

\section{Introduction}\label{introduction}

Many community broadband projects require a long-range wireless link
for their connection to the Internet. The equipment of choice is
currently based on wifi operating in the unlicensed 2.4 or 5GHz
spectra. This is cheap: the equipment for a long-distance
point-to-point link costs under \pounds 400 and can provide
bi-directional throughput of about 50Mb/s. While this provides an
improvement for the many rural communities that are served by long
copper telephone lines, 50Mb/s is no longer adequate for a community
of, say, 50 residences. Technically, there is no problem in getting
more bandwidth in one of the licenced spectra, but the equipment is
more expensive and the additional cost of the licence makes this
option unaffordable for small communities.

Recently some new equipment operating at 24GHz has come on the market
from Ubiquiti\footnote{\href{http://www.ubnt.com/airfiber}{\url{http://www.ubnt.com/airfiber}}}.
This is advertised as offering 1.4Gb/s at up to 13km. Although not as
cheap (a point-to-point link costs about \pounds 3,000) it might present
an opportunity for some rural communities to upgrade their service to
be competitive with the current fibre based offerings in the UK,
assuming they can find an internet connection with that bandwidth. A
5GHz variant is also produced by Ubiquiti claiming comparable speeds
at upwards of 50km.

With this in mind, the Scottish Government's Demonstrating
Digital programme provided the University of Edinburgh with funds to
test this equipment ``in the wild''. Of course, there is ample evidence
that the equipment works, but most of the evidence we have is
from installations in urban areas over short distances. How will
it perform over longer distances in West Highland weather? And what
are the practical problems faced by communities who want to install it?

At the request of the Scottish Government we also evaluated an
unrelated technology for use in urban areas -- free-space optics. This
means signalling by laser through the air instead of through
fibre-optic cabling. The main question here was, since these devices
operate in the visible spectrum, how well do they operate in reduced
visibility conditions? Do they operate \emph{well enough} for use in
this climate over short distances. Again it is a question of cost. The
equipment is expensive (some \pounds 8-15k per link) and requires
careful mounting and alignment, but compared to the civil engineering
cost of running fibre in urban areas, or leasing it, it may be worth
it.

\input{24GHz}
\unput{5GHz}
\input{lasers}
\end{document}
